---
title: "Getting started with rcore"
author:
  - name: Upendra Bhattarai
    affiliation: Harvard Chan Bioinformatics Core
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Getting started with rcore}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 4,
  dpi = 100,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

`rcore` provides a toolkit for analysts doing downstream analysis of
[nf-core](https://nf-co.re/) pipeline outputs.  It bundles:

- **R Markdown templates** for RNA-seq, single-cell, ChIP-seq, spatial
  transcriptomics, methylation, and multi-omics reporting.
- A **colorblind-friendly colour palette** based on teal and rose.
- A **ggplot2 theme** (`theme_rcore`) consistent with the report CSS.
- **Helper functions** for deploying templates, installing dependencies, and
  personalising reports with your own author name and contact details.

`rcore` was originally developed at the Harvard Chan Bioinformatics Core
and is now maintained as an open community package.

# Installation

```{r install, eval=FALSE}
if (!requireNamespace("remotes", quietly = TRUE))
    install.packages("remotes")
remotes::install_github("upendrabhattarai/rcore")
```

After installing, load the package:

```{r load}
library(rcore)
library(ggplot2)
```

# Personalising your reports

Run `rcore_setup()` once after installing the package.  Your name and
e-mail are stored in a user-level config file and will appear in every
report you deploy from that point on.

```{r setup-config, eval=FALSE}
rcore_setup(
  author   = "Your Name",
  email    = "you@example.com",
  org_abbr = "mylab"
)
```

Check the stored settings at any time:

```{r get-config, eval=FALSE}
rcore_get_config()
```

# Colorblind-friendly palette

`rcore` ships a 16-colour palette built around teal and rose — designed
to be distinguishable for the most common forms of colour-blindness
(deuteranopia/protanopia) while remaining visually appealing.

## Viewing the colours

```{r palette-names}
list_cb_friendly_cols()
```

## Using with ggplot2

The palette integrates directly with ggplot2 via
`scale_color_cb_friendly()` and `scale_fill_cb_friendly()`:

```{r palette-scatter, fig.cap="Diamond price vs. carat coloured by clarity using the rcore palette."}
set.seed(42)
dsamp <- diamonds[sample(nrow(diamonds), 800L), ]

ggplot(dsamp, aes(carat, price, colour = cut)) +
  geom_point(alpha = 0.7, size = 1.8) +
  scale_color_cb_friendly() +
  theme_rcore() +
  labs(title = "Diamond price vs. carat",
       subtitle = "Coloured by cut — rcore palette",
       x = "Carat", y = "Price (USD)")
```

You can also set the palette globally for the whole R session:

```{r global-scale, eval=FALSE}
theme_set(theme_rcore(base_size = 13))
scale_colour_discrete <- function(...) scale_color_cb_friendly(...)
scale_fill_discrete   <- function(...) scale_fill_cb_friendly(...)
```

## Available palette variants

```{r palette-variants, fig.cap="Heatmap showing the teal\u2013white\u2013rose diverging palette."}
pal_fn <- cb_friendly_pal("heatmap")
image(matrix(seq_len(20L)), col = pal_fn(20L), axes = FALSE,
      main = "heatmap palette (rose \u2013 white \u2013 teal)")
```

# ggplot2 theme

`theme_rcore()` is a clean, minimal theme with teal accent colours.
It can be used standalone or combined with any `scale_*` function.

```{r theme-demo, fig.cap="Bar chart using theme_rcore with the built-in teal palette."}
ggplot(mpg, aes(class, fill = drv)) +
  geom_bar(position = "dodge") +
  scale_fill_cb_friendly(palette = "cool") +
  theme_rcore() +
  labs(title = "Vehicle class by drive type",
       x = "Vehicle class", y = "Count", fill = "Drive")
```

# Deploying report templates

## Set up a new project

`rcore_templates()` copies Rmd reports and supporting files into a folder
of your choice.  The `author:` field in every deployed Rmd is
automatically updated to the name you stored with `rcore_setup()`.

```{r templates-base}
tmp_dir <- withr::local_tempdir()
rcore_templates(type = "base", outpath = tmp_dir)
fs::dir_ls(tmp_dir, recurse = TRUE, all = TRUE)
```

## Deploy analysis templates

Supported types are `"rnaseq"`, `"singlecell"`, `"singlecell_delux"`,
`"spatial"`, `"peakseq"`, and `"multiomics"`.

```{r templates-rnaseq}
reports_dir <- fs::path_join(c(tmp_dir, "reports"))
rcore_templates(type = "rnaseq", outpath = reports_dir)
fs::dir_ls(reports_dir, all = TRUE)
```

# Installing analysis dependencies

Use `install_rcore_deps()` to install all packages needed for a specific
workflow.  The function handles CRAN, Bioconductor, and GitHub-only
packages automatically.

```{r install-deps, eval=FALSE}
# Install only what RNA-seq templates need
install_rcore_deps("rnaseq")

# Install everything (takes a while)
install_rcore_deps("all")
```

See what each workflow needs before installing:

```{r list-deps}
list_rcore_deps()
```

# Injecting the CSS theme into a report

Add this chunk near the top of any `.Rmd` file to activate the rcore
HTML styling (dot-grid background, section cards, professional footer):

````
```{r rcore-css, echo=FALSE, results='asis'}
rcore::rcore_inject_css()
```
````

# Session information

```{r sessioninfo}
sessionInfo()
```
