---
title: "Celltype composition with Propeller"
author: "Harvard Chan Bioinformatics Core"
date: "`r Sys.Date()`"
output:
   html_document:
      code_folding: hide
      df_print: paged
      highlight: pygments
      number_sections: true
      self_contained: true
      theme: cosmo
      toc: true
      toc_depth: 3
      toc_float:
         collapsed: true
         smooth_scroll: true
params:
   project_file: ../information.R
---

```{r, message=FALSE, warning=FALSE}
# This set up the working directory to this file so all files can be found
library(rstudioapi)
library(tidyverse)
setwd(fs::path_dir(getSourceEditorContext()$path))
# NOTE: This code will check version, this is our recommendation, it may work
#.      other versions
stopifnot(R.version$major>= 4) # requires R4
if (compareVersion(R.version$minor,"3.1")<0) warning("We recommend >= R4.3.1") 
stopifnot(compareVersion(as.character(BiocManager::version()), "3.18")>=0)
stopifnot(compareVersion(as.character(packageVersion("Seurat")), "5.0.0")>=0)
```

This code is in this ![](https://img.shields.io/badge/status-draft-grey) revision.


```{r bcbio-css, echo=FALSE, results='asis'}
bcbioR::bcbio_inject_css()
```

```{r setup, cache=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# Load libraries
library(knitr)
library(rmarkdown)
library(Seurat)
library(SummarizedExperiment)
library(SingleCellExperiment)
library(DEGreport)
library(pheatmap)
library(tidyverse)
library(data.table)
library(DT)
library(patchwork)
library(future)
library(speckle)
library(bcbioR)

# ---- Global ggplot2 theme & colour scales ----
ggplot2::theme_set(bcbioR::theme_bcbio(base_size = 13))
scale_colour_discrete <- function(...) bcbioR::scale_color_cb_friendly(...)
scale_fill_discrete   <- function(...) bcbioR::scale_fill_cb_friendly(...)

# ---- Global knitr options ----
knitr::opts_chunk$set(
  cache      = FALSE,
  cache.lazy = FALSE,
  dev        = c("png", "pdf"),
  dpi        = 150,
  error      = TRUE,
  echo       = TRUE,
  fig.align  = "center",
  fig.height = 5,
  fig.retina = 2,
  fig.width  = 9.6,
  message    = FALSE,
  tidy       = FALSE,
  warning    = FALSE
)

if (future::supportsMulticore()) {
  future::plan(future::multicore)
} else {
  future::plan(future::multisession)
}
```

```{r sanitize_datatable}
sanitize_datatable = function(df, ...) {
 # remove dashes which cause wrapping
 DT::datatable(df, ..., rownames=gsub("-", "_", rownames(df)),
                   colnames=gsub("-", "_", colnames(df)),
                 options = list(scrollX = TRUE,...))
}
```

## Load object and subset
Read in the full dataset Seurat object and subset to keep only cells from the conditions of interest (cold7 vs TN).

```{r load-data}
seurat <- readRDS("data.rds")

subset_seurat <-  subset(x = seurat, subset= condition %in% c("Ctrl","Treat"))

Idents(object = subset_seurat) <- "celltype"

DefaultAssay(subset_seurat) <- "RNA"

# Create metadata df and factor celltype
meta <- subset_seurat@meta.data
meta$celltype <- factor(meta$celltype)
```

## Check count numbers of cells

```{r cellnumbers}
meta$condition_sample <- paste0(meta$condition, "_", meta$sample)
table(meta$condition_sample, meta$celltype)
```

## Run propeller

```{r run-propeller}
# working code; takes into account cell numbers
# NOTE adapt with your variables of interest
propres <- propeller(subset_seurat, sample=subset_seurat$sample,
                     clusters = subset_seurat$celltype,
                     group = subset_seurat$condition)

sanitize_datatable(propres)

```

## Proportions of cells table

To understand discordance with sccomp results, let's evaluate proportions and variability within group. 


```{r}
props <- getTransformedProps(meta$celltype, 
                                 meta$condition_sample, 
                                 transform="logit")

props$Proportions
```

```{r extracode, eval=FALSE}
meta_propeller<- meta %>% 
  group_by(condition, sample) %>%
  distinct(sample)

designAS <- model.matrix(~ meta_propeller$condition) 
colnames(designAS) <- c("Ctrl", "Treat")

fit <- lmFit(props$TransformedProps, designAS)
fit <- eBayes(fit, robust=TRUE)

topTable(fit)
```
# Resources

* [Speckle vignette](http://127.0.0.1:29356/library/speckle/doc/speckle.html#finding-significant-differences-in-cell-type-proportions-using-propeller)
* [Application to mouse PBMC scRNA](https://phipsonlab.github.io/propeller-paper-analysis/RealDataAnalysis.html#Young_vs_Old)

# Conclusion

# R session

List and version of tools used for the QC report generation.

```{r}
sessionInfo()
```


